---
title: "Gov 50 Final Project"
author: "Ava Berg"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---   
   

  
## Project thoughts

I am interested in exploring data related to political participation among undergraduate students, possibly comparing in-state and out-of-state. It could be useful to understand voting patterns across neighborhoods in Madison as well, analyzing different wards. I also would be interested in possibly analyzing IMF voting patterns, comparing rotating members and permanent members. 


## Project Milestone #2

For my final project I would like to find an answer for the following research question: how have prospects for gender equality progressed in response to the electoral competition and changes in national leadership in Algeria from 1989 to present? Decades after the Algerian War, initiatives for advancing the status of Women in Algeria have waned, in response to political instability over the last 3-and-a-half decades. Algeria has experienced several violent coups and wars during the state’s history, in which leadership and electoral party dominance has been in a state of flux.  I would like to study the outcomes for women following changes in electoral dominance of the state’s leading parties, as well as changes in leadership of the head of state. According to my hypothesis, political instability at the highest level and electoral competition have weakened outcomes for gender equality in Algeria. If I were to observe smaller margins in electoral victories and instances of regime change alongside diminished progress for women this would prove my hypothesis correct. To study gender equality outcomes I will use World Bank data on female labor force participation, alongside qualitative research on the status of family protection laws and other relevant legislation. In regards to the explanatory variables, I will study electoral competition via the margin of victory for the President, as well as the dominant party in the Algerian Parliament.

## Code

```{r}

my_data <- read.csv("MyData.csv.csv")

head(my_data)



```

## Project Milestone #3

```{r}

library(dplyr)
library(tidyr)

```

```{r}

long_data <- my_data |>
  pivot_longer(
    cols = `X1994`:`X2025`,
    names_to = "year",
    values_to = "labor_participation"
  ) |>
  select(-Indicator.Code) |>
    group_by(Country.Name, year) |>
  summarize(
    labor_participation = mean(labor_participation, na.rm = TRUE),
    .groups = "drop"
  ) 

long_data

```



```{r}

library(ggplot2)

```


```{r}

long_data |>
  filter(Country.Name == "Algeria") |>
  mutate(year = as.numeric(gsub("X", "", year)))|>
  ggplot(aes(x = year, y = labor_participation)) +
  geom_line() +
  labs(
    title = "Female Labor Force Participation in Algeria",
    x = "Year",
    y = "Participation"
  ) 

```

Final Project Milestone #4:

Your article should contain one visualization and one analysis that attempts to answer your research question. There does not need to be a long discussion, but the results should be presented in either a second visualization or a nicely formatted table.





```{r}

library(jsonlite)

# Fetch the data
democracy_data <- read.csv("https://ourworldindata.org/grapher/electoral-democracy-index.csv?v=1&csvType=full&useColumnShortNames=false")

# Fetch the metadata
metadata <- fromJSON("https://ourworldindata.org/grapher/electoral-democracy-index.metadata.json?v=1&csvType=full&useColumnShortNames=false")

head(democracy_data)

```


```{r}

library(tidyverse)


algeria_data <- democracy_data |>
  filter(Entity == "Algeria") |>
  select(Entity, Year, Electoral.Democracy.Index)

algeria_data

algeria_data1 <- long_data |>
  filter(Country.Name == "Algeria") |>
  mutate(year = as.numeric(gsub("X", "", year))) 


algeria_data1 <- algeria_data1 |>
  rename(Entity = Country.Name, Year = year)

algeria_data1


```

```{r}

merged_data <- inner_join(algeria_data1, algeria_data, by = c("Entity", "Year"))

```


```{r}
  ggplot(merged_data, aes(x = Electoral.Democracy.Index, y = labor_participation)) +
  geom_line() +
  labs(
    x = "Electoral Demcracy Score",
    y = "Women's Labor Participation",
    title = "Electoral Competition and Female Empowerment in Algeria"
  ) 



```

   
Analysis:

In this visualization I compared female labor force participation from the World Bank, alongside Algeria's electoral democracy score. The electoral democracy score was sourced from the Electoral Democracy Index and analyzing the competitiveness of Algeria's democratic system, beginning in the year 1994. When comparing these two variables, it appears there is not evidence of a correlation between democracy competitiveness and female empowerment. Confounding factors, such as microeconomic conditions, are likely to have a larger impact on the status of women in Algeria, as opposed to electoral competition.   

 

