---
title: "Gov 50 Final Project"
author: "Ava Berg"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false 
---   
  
   Political scientists have long studied what factors promote democracy. In 2008, Valentine Mogham proposed that women’s empowerment and equality were vital to sustaining a democracy. Thus, I would like to analyze how gender equality has progressed alongside democratic backsliding.I will perform a longitudinal observational study to find an answer to the following research question: how have prospects for gender equality progressed in response to the electoral competition in Algeria from 1994 to present? According to my hypothesis, democratic backsliding is associated with worse outcomes for female empowerment in Algeria. These findings are important for forming a greater understanding of how to promote and sustain democracy. 
   To evaluate this hypothesis I will use two key variables. The independent variable is the strength of Algeria’s democracy, which will be measured via the data from the Electoral Democracy Index. The Electoral Democracy Index estimates the extent to which political leaders are elected under comprehensive voting rights in free and fair elections, where freedoms of association and expression are guaranteed on a scale of 0 to 1. To measure my dependent variable, gender equality outcomes, I will use World Bank data on female labor force participation. Female labor force participation intersects with various aspects of female empowerment, such as economic independence and the ability to work outside of the home.In the second code chunk a chart visualizes these two variables.  
  I expect to observe lower electoral democracy index scores, alongside decreasing female labor force participation. I will run a regression in which the dependent/outcome variable is female empowerment and the independent/explanatory variable is the strength of Algeria’s democracy. To reject my hypothesis I would expect to see a positive correlation from the regression coefficient. Additionally, if there is not sufficient evidence to reject the null hypothesis that there is not a relationship between democracy and female empowerment. On the other hand, the alternative hypothesis would suggest that higher electoral democracy index scores were correlated with higher scores for female labor force participation. 



```{r}

library(dplyr)
library(tidyr)
library(ggplot2)


my_data <- read.csv("MyData.csv.csv")


long_data <- my_data |>
  pivot_longer(
    cols = `X1994`:`X2025`,
    names_to = "year",
    values_to = "labor_participation"
  ) |>
  select(-Indicator.Code) |>
    group_by(Country.Name, year) |>
  summarize(
    labor_participation = mean(labor_participation, na.rm = TRUE),
    .groups = "drop"
  ) 



```



```{r}

library(ggplot2)
library(dplyr)

long_data |>
  filter(Country.Name == "Algeria") |>
  mutate(year = as.numeric(gsub("X", "", year)))|>
  ggplot(aes(x = year, y = labor_participation)) +
  geom_line() +
  labs(
    title = "Female Labor Force Participation in Algeria",
    x = "Year",
    y = "Participation"
  ) 


```




```{r}

library(jsonlite)


democracy_data <- read.csv("https://ourworldindata.org/grapher/electoral-democracy-index.csv?v=1&csvType=full&useColumnShortNames=false")
metadata <- fromJSON("https://ourworldindata.org/grapher/electoral-democracy-index.metadata.json?v=1&csvType=full&useColumnShortNames=false")

algeria_data <- democracy_data |>
  filter(Entity == "Algeria") |>
  select(Entity, Year, Electoral.Democracy.Index)

algeria_data1 <- long_data |>
  filter(Country.Name == "Algeria") |>
  mutate(year = as.numeric(gsub("X", "", year))) 


algeria_data1 <- algeria_data1 |>
  rename(Entity = Country.Name, Year = year)




```

```{r}

merged_data <- inner_join(algeria_data1, algeria_data, by = c("Entity", "Year"))

```


```{r}

  
ggplot(merged_data, aes(x = Electoral.Democracy.Index, y = labor_participation)) +
  geom_point() +
  labs(
    x = "Electoral Demcracy Index Score",
    y = "Women's Labor Participation",
    title = "Electoral Democracy and Female Empowerment in Algeria",
    subtitle = "1994-2025"
  ) +
  geom_smooth(method = "lm", se = FALSE, color = "indianred1", size = 1.5)

```
  
   
```{r}

summary(lm(labor_participation ~ Electoral.Democracy.Index, data = merged_data))


```
 

  After the regression analysis, I found the following coefficients: (Intercept): 14.050, Electoral.Democracy.Index: 4.398. Therefore, when the democracy score is zero, we could expect female labor participation to be approximately 14.1%. The slope of 4.398 indicates that when democracy increases from 0 to 1, we can expect female labor force participation to increase by 4.398. 
	However, it would appear that democracy is not correlated with female labor force participation in Algeria. The p-value of 0.611 is less than 0.05, so we are unable to reject the null hypothesis.  Additionally, the R² statistic is 0.008748, which means that the regression model can only explain 0.9% of the variation in female labor force participation. 
	
	
   Since the null hypothesis has not been rejected and the r-squared statistic is extremely low, democracy cannot be attributed as a cause of improved gender empowerment outcomes in the context of Algeria 1994 onward. Instead, it is possible that cultural norms surrounding gender, as well as economic conditions may play a larger role in shaping female labor force participation. There are also other methods to study female empowerment, such as female representation in government which would offer a more complete picture of gender equality in Algeria. If I had more time and money I would include female representation in Algeria’s national and local governments, as well as similar regressions for Morocco and Tunisia. This is because these countries have also had revolutionary pasts and are part of the Maghreb as well.
 

